RIScatter: Unifying Backscatter Communication and Reconfigurable Intelligent Surface

### 2022-12-18 jsac_221218.tex
:construction: Add JSAC version [in progress]

\documentclass[journal]{IEEEtran}
% \documentclass[journal,12pt,onecolumn,draftclsnofoot]{IEEEtran}

\usepackage[table]{xcolor}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage[american]{circuitikz}
\usepackage{cite}
\usepackage{fixmath}
\usepackage[acronym]{glossaries-extra}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage[short]{optidef}
\usepackage{pgfplots}
\usepackage{ragged2e}
\usepackage[subtle]{savetrees}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{tabularx}
\usepackage{tikz}
% dark mode
% \usepackage{xcolor} \pagecolor[rgb]{0,0,0} \color[rgb]{1,1,1}

@@

\begin{section}{Introduction}
	\begin{subsection}{Fundamentals}
		\IEEEPARstart{F}{uture} wireless network is envisioned to provide high throughput, uniform coverage, pervasive connectivity, heterogeneous control, and cognitive intelligence for trillions of portable devices.
		As a mature low-power communication technique, \gls{bc} separates a conventional transmitter into a \gls{rf} carrier emitter with power-hungry elements (e.g., synthesizer and amplifier) and an information-bearing node with power-efficient components (e.g., harvester and modulator) \cite{Boyer2014}.
		In particular, the node harvests energy from impinging wave and embeds information over scattered signal, and the backscatter reader can be either co-located or separated with the carrier emitter, known as \gls{mbc} in Fig.~\subref*{fg:mbc} and \gls{bbc} in Fig.~\subref*{fg:bbc}.
		Relevant applications such as \gls{rfid} \cite{Dobkin2012,Landt2005} and passive sensor network \cite{Vannucci2008,Assimonis2016} have been extensively researched, standardized, and commercialized in the era of \gls{ioe}.
		However, conventional \gls{bc} nodes only respond when externally inquired by a nearby reader.
		To tackle this, \cite{Liu2013b} proposed \gls{ambc} where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connection in between.
		As shown in Fig.~\subref*{fg:ambc}, it eliminates the need of dedicated power supply, carrier emitter and frequency spectrum for \gls{bc}, but the decoding is subject to a strong interference from the primary (legacy) link.
		Cooperative \gls{ambc} was then proposed where a co-located receiver decodes both active primary and passive backscatter links, and the rate and error performance are evaluated for various detection schemes \cite{Yang2018}.
		The concept was further refined by \cite{Liang2020} as \gls{sr} in Fig.~\subref*{fg:sr} that cognitively incorporates \gls{ambc} with existing systems.
		In a \gls{sr} system, the active transmitter generates \gls{rf} wave carrying primary information, the passive node creates additional propagation path and ride its information over scattered signal, while the cooperative receiver jointly, individually, or sequentially decodes both links.
		Those \gls{bc} applications employ scatter nodes as pure information sources, and the active primary transmission (if it exists) can be influenced by the randomness of backscatter modulation.
		On the other hand, \gls{ris} in Fig.~\ref{fg:ris} is a smart planar signal reflector that consists of numerous compact passive elements with adjustable amplitude and phase responses, whose reflection pattern over time is deterministic and adaptively optimized based on \gls{csi}.
		\gls{ris} smartly reflects the surrounding \gls{rf} waves to customize the wireless propagation environment for signal enhancement, interference suppression, scattering enrichment and non-line-of-sight bypassing \cite{Wu2021b}.
		In summary, \gls{mbc}/\gls{bbc} (resp. \gls{ris}) involves a \emph{single} passive backscatter (resp. active primary) link where the passive device modulates information (resp. enhances propagation) and the active device generates carrier wave (resp. modulated signal).
		On the other hand, \gls{ambc} (resp. \gls{sr}) involves \emph{coexisting} active primary and passive backscatter links in a competitive (resp. collaborative) relationship.
		In \gls{ambc}, the backscatter link only needs to know the basic information (e.g., operating frequency) of the primary link, while the primary link may be completely unaware of the backscatter link.
		In \gls{sr}, relevant \gls{csi} is exploited at the co-located receiver for cooperative decoding, and may also be exploited at the primary transmitter for cooperative active beamforming.
		A comparison of scattering applications is summarized in Table~\ref{tb:scattering_applications}.

		\begin{figure}[!t]
			\centering
			\subfloat[\gls{mbc}]{
				\resizebox{0.48\linewidth}{!}{
					\input{assets/illustration/mbc.tex}
				}
				\label{fg:mbc}
			}
			\subfloat[\gls{bbc}]{
				\resizebox{0.48\linewidth}{!}{
					\input{assets/illustration/bbc.tex}
				}
				\label{fg:bbc}
			}
			\\
			\subfloat[\gls{ambc}]{
				\resizebox{0.48\linewidth}{!}{
					\input{assets/illustration/ambc.tex}
				}
				\label{fg:ambc}
			}
			\subfloat[\gls{sr}]{
				\resizebox{0.48\linewidth}{!}{
					\input{assets/illustration/sr.tex}
				}
				\label{fg:sr}
			}
			\\
			\subfloat[\gls{ris}]{
				\resizebox{0.48\linewidth}{!}{
					\input{assets/illustration/ris.tex}
				}
				\label{fg:ris}
			}
			\subfloat[RIScatter]{
				\resizebox{0.48\linewidth}{!}{
					\input{assets/illustration/riscatter.tex}
				}
				\label{fg:riscatter}
			}
			\caption{
				Illustration of scattering applications.
				For each, the blue flow(s) constitute the active primary link while the magenta/green flow denotes the passive backscatter link.
			}
			\label{fg:scatter_illustration}
		\end{figure}
	\end{subsection}

	\begin{table*}[!t]
		\caption{Comparison of Scattering Applications}
		\label{tb:scattering_applications}
		\rowcolors{2}{gray!25}{white}
		\renewcommand{\arraystretch}{1.4}
		\begin{tabularx}{\textwidth}{CCCCCC}
			\toprule
			\hiderowcolors
			                                                        & \gls{mbc}/\gls{bbc} & \gls{ambc}                              & \gls{sr}                                              & \gls{ris}           & RIScatter                                                               \\ \midrule
			\showrowcolors
			% Coexisting link(s)                                              & 1                   & 2 (competitive)             & 2 (collaborative)                                     & 1                   & 2 (collaborative)                                                       \\
			Information link(s)                                     & Passive backscatter & Active-passive coexisting (competitive) & Active-passive coexisting (collaborative)             & Active primary      & Active-passive coexisting (collaborative)                               \\
			% Primary/Backscatter modulation                                  & No/Yes              & Yes/Yes                     & Yes/Yes                                               & Yes/No              & Yes/Yes                                                                 \\
			Primary impact on backscatter detection                 & Carrier             & Multiplicative interference             & Spreading code or multiplicative interference         & ---                 & Energy uncertainty                                                      \\
			Backscatter impact on primary detection                 & ---                 & Multiplicative interference             & \gls{csi} uncertainty or multiplicative interference  & Passive beamforming & Dynamic passive beamforming                                             \\
			Collaborative devices based on \gls{csi}                & ---                 & ---                                     & Co-located receiver (potentially primary transmitter) & ---                 & Primary transmitter, scatter nodes, and co-located receiver             \\
			Cooperative decoding strategy                           & ---                 & No                                      & Usually \gls{sic} from primary to backscatter         & ---                 & No \gls{sic}, backscatter detection as part of primary channel training \\
			Primary detection                                       & ---                 & Semi-coherent                           & Semi-coherent                                         & Coherent            & Coherent                                                                \\
			Backscatter detection                                   & Coherent            & Semi-coherent                           & Coherent                                              & ---                 & Semi-coherent                                                           \\
			Reflection pattern over time depends on                 & Information source  & Information source                      & Information source                                    & \gls{csi}           & Information source, \gls{csi}, and \gls{qos}                            \\
			Reflection state (i.e., backscatter input) distribution & Equiprobable        & Equiprobable                            & Equiprobable or Gaussian                              & Degenerate          & Flexible                                                                \\
			Load-switching speed                                    & Fast                & Fast or slow                            & Very slow                                             & Quasi-static        & Arbitrary                                                               \\ \bottomrule
		\end{tabularx}
	\end{table*}

	\begin{subsection}{Related Works}
		Similar to \gls{cr}, active-passive coexisting networks can be classified into commensal, parasitic, and competitive relationships, whose instantaneous rates, power schemes, and outage probabilities were acquired in \cite{Guo2019b,Ding2020}.
		% To evaluate the performance of cooperative receivers, \cite{Yang2018} derived the bit error rates of \gls{ml}, linear and \gls{sic} detectors for flat fading channels, and proposed a low-complexity detector for frequency-selective fading channels.
		% However, it only considered the case where the primary and backscatter symbols of the same period are perfectly aligned in time.
		Joint \gls{ml} decoding provides the best error performance but requires prohibitive computational complexity especially for high-order modulations \cite{Yang2018}.
		Indeed, active-passive coexisting networks have three special and important properties:
		1) primary and backscatter symbols are superimposed by \emph{double modulation} (i.e., multiplication coding);
		2) backscatter signal strength is much weaker than primary due to the \emph{double fading} effect;
		3) backscatter \emph{symbol period} is longer than primary due to the load switching constraint.
		% ! double modulation, double fading, symbol period difference
		The second property motivated \cite{Long2020a,Liang2020,Guo2019b,Ding2020,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Yang2018,Han2021,Zhang2022} to view \gls{sr} as a multiplicative \gls{noma} and perform \gls{sic} from primary to backscatter link.
		During primary decoding, the randomness of backscatter modulation can be modelled as channel uncertainty or multiplicative interference when the backscatter-over-primary symbol period ratio $N$ is large or small\footnote{For state-of-the-art low-power \gls{rfid} tags, the load-switching speed usually varies between \num{100}s of \unit{\kHz} to \num{10}s of \unit{\MHz} \cite{Torres2021}, corresponding to a typical $N$ between \num{10} and \num{1000}.}.
		Such a \gls{sic} receiver is operation-intensive and \gls{csi}-sensitive because it involves $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} to decode each backscatter symbol.
		When $N$ is sufficiently large (i.e., commensal \gls{sr}), the average primary achievable rate under semi-coherent detection\footnote{In this paper, semi-coherent detection refers to the primary/backscatter decoding based on known \gls{csi} and unknown backscatter/primary symbols.} asymptotically approaches its coherent counterpart, and both links may be decoded in an interference-free manner \cite{Long2020a}.
		On the other hand, efficient backscatter multiple access design remains an open issue for \gls{bc}.
		\cite{Xu2021a} proposed a \gls{noma}-based \gls{sr} where the \gls{sic} order depends on the backscatter channel strength, but the performance deteriorates fast as the number of nodes increases.
		\gls{tdma}-based \gls{sr} was also considered in \cite{Yang2021a} where each node transmits information during its dedicated slot and harvests energy otherwise.
		It enables adaptive transmission time and reflection ratio design but introduces extra coordination cost.
		\cite{Vougioukas2019} modifies the load-switching speed to shift the scattered signal to the desired frequency band.
		This enables backscatter \gls{fdma} at the cost of larger bandwidth and higher power consumption at the scatter nodes.
		To reduce the coordination cost, \cite{Han2021} proposed a random code-assisted multiple access for \gls{sr} and evaluated the asymptotic \gls{sinr} using random matrix theory.
		This code-domain solution suffers from the near-far problem and imperfect synchronization, both being severe in \gls{bc}.

		Conventional \gls{ris} design with fixed reflection coefficients during each channel block has been extensively studied in communication, sensing, and power literatures \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
		On the other hand, dynamic \gls{ris} with independent reflection patterns in different time slots has gained recent attentions in multi-user and multi-purpose wireless networks.
		The concept was first proposed in \cite{Yang2020} to fine-tune the \gls{ofdm} resource blocks, then extended to the downlink power and uplink information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
		It introduces artificial channel diversity and enables flexible resource allocation, but misses the opportunity to encode its own message.
		\gls{ris} can also be used as a transmitter when placed in the near field of a carrier emitter, and prototypes have been implemented for \gls{psk} \cite{Tang2019a} and \gls{qam} \cite{Dai2020a}.
		From an information-theoretic perspective, \cite{Karasik2020} reported that using \gls{ris} as a naive passive beamformer to maximize the receive \gls{snr} is generally rate-suboptimal for finite-input constellations.
		Instead, the capacity of \gls{ris}-aided channel is achieved by joint transmitter-\gls{ris} encoding, and \gls{sic}-based \gls{sr} can outperform passive beamforming at high \gls{snr}.
		It inspired \gls{ris}-empowered \gls{bc} \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to combine passive beamforming and backscatter modulation in the overall reflection pattern.
		In particular, \emph{symbol level precoding} maps the information symbols to the optimized \gls{ris} coefficient sets \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes the information symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between the reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} employs dedicated reflection elements (resp. information elements) for passive beamforming (resp. backscatter modulation) \cite{Hu2021a}.
		However, those joint designs incur advanced hardware architecture and high optimization complexity.
		Relevant literatures also consider either \emph{Gaussian} codebook \cite{Guo2019b,Ding2020,Long2020a,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Hu2021b} that is impractical for low-power nodes, or finite \emph{equiprobable} inputs \cite{Yang2018,Liang2020,Han2021,Zhang2022,Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} that does not fully exploit the \gls{csi} and signal properties of active-passive coexisting networks.
	\end{subsection}

	\begin{subsection}{Contributions}
		\emph{First,} we propose RIScatter as a novel scatter protocol that unifies \gls{bc} and \gls{ris} by adaptive reflection state (i.e., backscatter input) distribution design.
		As shown in Fig.~\subref*{fg:riscatter}, one or more passive RIScatter nodes ride over an active transmission and smartly scatters the incident wave to simultaneously engineer the primary channel and modulate its own information, while a co-located receiver cooperatively decodes both coexisting links.
		The reflection pattern over time is neither fully random like \gls{bc} nor fully deterministic like \gls{ris}, but semi-randomly chosen with the guidance of reflection state distribution, whose design accommodates the impact of information source, \gls{csi}, and \gls{qos} of both links.
		Such an adaptive backscatter channel coding boils down to the degenerate distribution of \gls{ris} when the primary link is prioritized, and outperforms the uniform distribution of line-coded \gls{bc} (by accounting the \gls{csi}) when the backscatter link is prioritized.
		Besides, when multiple RIScatter nodes are co-located, joint encoding can further increase the total backscatter rate.
		RIScatter exploits the transmitter-scatter-receiver path for simultaneous backscatter modulation and dynamic passive beamforming, and can smoothly transition in between under different \gls{qos}.
		However, such a novel scatter protocol encounters two major challenges, namely the cooperative receiver design and the adaptive backscatter input distribution design.
		This is the first paper to unify \gls{bc} and \gls{ris} from the perspective of reflection state distribution.

@@

		\emph{Fourth,} we provide numerical results to demonstrate the benefits of RIScatter and proposed algorithms.
		The observations include:
		1) adaptive reflection state distribution design can flexibly transition between backscatter modulation and passive beamforming;
		2) when the primary link is prioritized, input distribution becomes degenerate and RIScatter nodes coincide with discrete \gls{ris};
		3) when the backscatter link is prioritized, adaptive RIScatter encoding achieves higher backscatter rate than conventional line-coded \gls{bc} with equiprobable inputs;
		4) co-located RIScatter nodes can further leverage total backscatter rate by joint encoding;
		5) the proposed receiver maintains the passive beamforming benefit and provides comparable backscatter rate to \gls{sic}-based \gls{sr}, with re-encoding costs reduced to $1/N$ and no re-precoding/cancellation;
		6) it also supports scatter nodes with faster load-switching speed for potentially higher throughput;
		7) \gls{pgd} active beamformer effectively increases the primary (resp. backscatter) rate by boosting the receive \gls{snr} (resp. widening the energy gap under different reflection states), and can smoothly transition in between under different \gls{qos};
		8) distribution-aware backscatter detectors provide higher backscatter rate than the conventional \gls{ml} detector.
	\end{subsection}

@@

			\def\svgwidth{0.7\columnwidth}

@@

		Due to double modulation, the primary equivalent channel is a function of backscatter symbol tuple\footnote{\eqref{eq:equivalent_channel_bc} and \eqref{eq:equivalent_channel_ris} are often used in \gls{bc} and \gls{ris} literatures, respectively.}

@@

				\boldsymbol{h}_{\text{E}}^\mathsf{H}(x_{\mathcal{K}})
				 & \triangleq \boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k \label{eq:equivalent_channel_bc}                    \\

>>>>

% \documentclass[journal]{IEEEtran}
\documentclass[journal,12pt,onecolumn,draftclsnofoot]{IEEEtran}

\usepackage[table]{xcolor}
\usepackage{adjustbox}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage[american]{circuitikz}
\usepackage{cite}
\usepackage{fixmath}
\usepackage[acronym]{glossaries-extra}
\usepackage{hyperref}
\usepackage{import}
\usepackage{mathtools}
\usepackage{microtype}
\usepackage[short]{optidef}
\usepackage{pgfplots}
\usepackage{ragged2e}
\usepackage[subtle]{savetrees}
\usepackage{setspace}
\usepackage{siunitx}
\usepackage{stfloats}
\usepackage[caption=false,font=footnotesize,subrefformat=parens,labelformat=parens]{subfig}
\usepackage{tabularx}
\usepackage{tikz}

% dark mode
% \usepackage{xcolor} \pagecolor[rgb]{0,0,0} \color[rgb]{1,1,1}

@@

\begin{section}{Introduction}
	% \begin{subsection}{Fundamentals}
	\IEEEPARstart{F}{uture} wireless network is envisioned to provide high throughput, uniform coverage, pervasive connectivity, heterogeneous control, and cognitive intelligence for trillions of portable devices.
	As a mature low-power communication technique, \gls{bc} separates a conventional transmitter into a \gls{rf} carrier emitter with power-hungry elements (e.g., synthesizer and amplifier) and an information-bearing node with power-efficient components (e.g., harvester and modulator) \cite{Boyer2014}.
	In particular, the node harvests energy from impinging wave and embeds information over scattered signal, and the backscatter reader can be either co-located or separated with the carrier emitter, known as \gls{mbc} in Fig.~\subref*{fg:mbc} and \gls{bbc} in Fig.~\subref*{fg:bbc}.
	Relevant applications such as \gls{rfid} \cite{Dobkin2012,Landt2005} and passive sensor network \cite{Vannucci2008,Assimonis2016} have been extensively researched, standardized, and commercialized in the era of \gls{ioe}.
	However, conventional \gls{bc} nodes only respond when externally inquired by a nearby reader.
	To tackle this, \cite{Liu2013b} proposed \gls{ambc} where battery-free nodes recycle ambient signals (e.g., radio, television and Wi-Fi) to harvest energy and establish connection in between.
	As shown in Fig.~\subref*{fg:ambc}, it eliminates the need of dedicated power supply, carrier emitter and frequency spectrum for \gls{bc}, but the decoding is subject to a strong interference from the primary (legacy) link.
	Cooperative \gls{ambc} was then proposed where a co-located receiver decodes both active primary and passive backscatter links, and the rate and error performance are evaluated for various detection schemes \cite{Yang2018}.
	The concept was further refined by \cite{Liang2020} as \gls{sr} in Fig.~\subref*{fg:sr} that cognitively incorporates \gls{ambc} with existing systems.
	In a \gls{sr} system, the active transmitter generates \gls{rf} wave carrying primary information, the passive node creates additional propagation path and ride its information over scattered signal, while the cooperative receiver jointly, individually, or sequentially decodes both links.
	Those \gls{bc} applications employ scatter nodes as pure information sources, and the active primary transmission (if it exists) can be influenced by the randomness of backscatter modulation.
	On the other hand, \gls{ris} in Fig.~\ref{fg:ris} is a smart planar signal reflector that consists of numerous compact passive elements with adjustable amplitude and phase responses, whose reflection pattern over time is deterministic and adaptively optimized based on \gls{csi}.
	\gls{ris} smartly reflects the surrounding \gls{rf} waves to customize the wireless propagation environment for signal enhancement, interference suppression, scattering enrichment and non-line-of-sight bypassing \cite{Wu2021b}.
	In summary, \gls{mbc}/\gls{bbc} (resp. \gls{ris}) involves a \emph{single} passive backscatter (resp. active primary) link where the passive device modulates information (resp. enhances propagation) and the active device generates carrier wave (resp. modulated signal).
	On the other hand, \gls{ambc} (resp. \gls{sr}) involves \emph{coexisting} active primary and passive backscatter links in a competitive (resp. collaborative) relationship.
	In \gls{ambc}, the backscatter link only needs to know the basic information (e.g., operating frequency) of the primary link, while the primary link may be completely unaware of the backscatter link.
	In \gls{sr}, relevant \gls{csi} is exploited at the co-located receiver for cooperative decoding, and may also be exploited at the primary transmitter for cooperative active beamforming.
	A comparison of scattering applications is summarized in Table~\ref{tb:scattering_applications}.

	\begin{figure}[!t]
		\centering
		\subfloat[\gls{mbc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/mbc.tex}
			}
			\label{fg:mbc}
		}
		\subfloat[\gls{bbc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/bbc.tex}
			}
			\label{fg:bbc}
		}
		\subfloat[\gls{ambc}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/ambc.tex}
			}
			\label{fg:ambc}
		}
		\\
		\subfloat[\gls{sr}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/sr.tex}
			}
			\label{fg:sr}
		}
		\subfloat[\gls{ris}]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/ris.tex}
			}
			\label{fg:ris}
		}
		\subfloat[RIScatter]{
			\resizebox{0.25\linewidth}{!}{
				\input{assets/illustration/riscatter.tex}
			}
			\label{fg:riscatter}
		}
		\caption{
			Illustration of scattering applications.
			For each, the blue flow(s) constitute the active primary link while the magenta/green flow denotes the passive backscatter link.
		}
		\label{fg:scatter_illustration}
	\end{figure}
	% \end{subsection}

	\begin{table*}[!t]
		\caption{Comparison of Scattering Applications}
		\label{tb:scattering_applications}
		\rowcolors{2}{gray!25}{white}
		\scriptsize
		\begin{tabularx}{\textwidth}{CCCCCC}
			\toprule
			\hiderowcolors
			                                                        & \gls{mbc}/\gls{bbc} & \gls{ambc}                              & \gls{sr}                                              & \gls{ris}           & RIScatter                                                               \\ \midrule
			\showrowcolors
			% Coexisting link(s)                                              & 1                   & 2 (competitive)             & 2 (collaborative)                                     & 1                   & 2 (collaborative)                                                       \\
			Information link(s)                                     & Passive backscatter & Active-passive coexisting (competitive) & Active-passive coexisting (collaborative)             & Active primary      & Active-passive coexisting (collaborative)                               \\
			% Primary/Backscatter modulation                                  & No/Yes              & Yes/Yes                     & Yes/Yes                                               & Yes/No              & Yes/Yes                                                                 \\
			Primary impact on backscatter detection                 & Carrier             & Multiplicative interference             & Spreading code or multiplicative interference         & ---                 & Energy uncertainty                                                      \\
			Backscatter impact on primary detection                 & ---                 & Multiplicative interference             & \gls{csi} uncertainty or multiplicative interference  & Passive beamforming & Dynamic passive beamforming                                             \\
			Collaborative devices based on \gls{csi}                & ---                 & ---                                     & Co-located receiver (potentially primary transmitter) & ---                 & Primary transmitter, scatter nodes, and co-located receiver             \\
			Cooperative decoding strategy                           & ---                 & No                                      & Usually \gls{sic} from primary to backscatter         & ---                 & No \gls{sic}, backscatter detection as part of primary channel training \\
			Primary detection                                       & ---                 & Semi-coherent                           & Semi-coherent                                         & Coherent            & Coherent                                                                \\
			Backscatter detection                                   & Coherent            & Semi-coherent                           & Coherent                                              & ---                 & Semi-coherent                                                           \\
			Reflection pattern over time depends on                 & Information source  & Information source                      & Information source                                    & \gls{csi}           & Information source, \gls{csi}, and \gls{qos}                            \\
			Reflection state (i.e., backscatter input) distribution & Equiprobable        & Equiprobable                            & Equiprobable or Gaussian                              & Degenerate          & Flexible                                                                \\
			Load-switching speed                                    & Fast                & Fast or slow                            & Very slow                                             & Quasi-static        & Arbitrary                                                               \\ \bottomrule
		\end{tabularx}
	\end{table*}

	% \begin{subsection}{Related Works}
	Similar to \gls{cr}, active-passive coexisting networks can be classified into commensal, parasitic, and competitive relationships, whose instantaneous rates, power schemes, and outage probabilities were acquired in \cite{Guo2019b,Ding2020}.
	% To evaluate the performance of cooperative receivers, \cite{Yang2018} derived the bit error rates of \gls{ml}, linear and \gls{sic} detectors for flat fading channels, and proposed a low-complexity detector for frequency-selective fading channels.
	% However, it only considered the case where the primary and backscatter symbols of the same period are perfectly aligned in time.
	Joint \gls{ml} decoding provides the best error performance but requires prohibitive computational complexity especially for high-order modulations \cite{Yang2018}.
	Indeed, active-passive coexisting networks have three special and important properties:
	1) primary and backscatter symbols are superimposed by \emph{double modulation} (i.e., multiplication coding);
	2) backscatter signal strength is much weaker than primary due to the \emph{double fading} effect;
	3) backscatter \emph{symbol period} is longer than primary due to the load switching constraint.
	% ! double modulation, double fading, symbol period difference
	The second property motivated \cite{Long2020a,Liang2020,Guo2019b,Ding2020,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Yang2018,Han2021,Zhang2022} to view \gls{sr} as a multiplicative \gls{noma} and perform \gls{sic} from primary to backscatter link.
	During primary decoding, the randomness of backscatter modulation can be modelled as channel uncertainty or multiplicative interference when the backscatter-over-primary symbol period ratio $N$ is large or small\footnote{For state-of-the-art low-power \gls{rfid} tags, the load-switching speed usually varies between \num{100}s of \unit{\kHz} to \num{10}s of \unit{\MHz} \cite{Torres2021}, corresponding to a typical $N$ between \num{10} and \num{1000}.}.
	Such a \gls{sic} receiver is operation-intensive and \gls{csi}-sensitive because it involves $N$ re-encoding, precoding, subtraction together with a time-domain \gls{mrc} to decode each backscatter symbol.
	When $N$ is sufficiently large (i.e., commensal \gls{sr}), the average primary achievable rate under semi-coherent detection\footnote{In this paper, semi-coherent detection refers to the primary/backscatter decoding based on known \gls{csi} and unknown backscatter/primary symbols.} asymptotically approaches its coherent counterpart, and both links may be decoded in an interference-free manner \cite{Long2020a}.
	On the other hand, efficient backscatter multiple access design remains an open issue for \gls{bc}.
	\cite{Xu2021a} proposed a \gls{noma}-based \gls{sr} where the \gls{sic} order depends on the backscatter channel strength, but the performance deteriorates fast as the number of nodes increases.
	\gls{tdma}-based \gls{sr} was also considered in \cite{Yang2021a} where each node transmits information during its dedicated slot and harvests energy otherwise.
	It enables adaptive transmission time and reflection ratio design but introduces extra coordination cost.
	\cite{Vougioukas2019} modifies the load-switching speed to shift the scattered signal to the desired frequency band.
	This enables backscatter \gls{fdma} at the cost of larger bandwidth and higher power consumption at the scatter nodes.
	To reduce the coordination cost, \cite{Han2021} proposed a random code-assisted multiple access for \gls{sr} and evaluated the asymptotic \gls{sinr} using random matrix theory.
	This code-domain solution suffers from the near-far problem and imperfect synchronization, both being severe in \gls{bc}.

	Conventional \gls{ris} design with fixed reflection coefficients during each channel block has been extensively studied in communication, sensing, and power literatures \cite{Wu2018,Zhang2019a,Lin2022,Liu2022,Feng2022,Zhao2022}.
	On the other hand, dynamic \gls{ris} with independent reflection patterns in different time slots has gained recent attentions in multi-user and multi-purpose wireless networks.
	The concept was first proposed in \cite{Yang2020} to fine-tune the \gls{ofdm} resource blocks, then extended to the downlink power and uplink information phases of \gls{wpcn} \cite{Wu2021,Wu2021d,Hua2022a}.
	It introduces artificial channel diversity and enables flexible resource allocation, but misses the opportunity to encode its own message.
	\gls{ris} can also be used as a transmitter when placed in the near field of a carrier emitter, and prototypes have been implemented for \gls{psk} \cite{Tang2019a} and \gls{qam} \cite{Dai2020a}.
	From an information-theoretic perspective, \cite{Karasik2020} reported that using \gls{ris} as a naive passive beamformer to maximize the receive \gls{snr} is generally rate-suboptimal for finite-input constellations.
	Instead, the capacity of \gls{ris}-aided channel is achieved by joint transmitter-\gls{ris} encoding, and \gls{sic}-based \gls{sr} can outperform passive beamforming at high \gls{snr}.
	It inspired \gls{ris}-empowered \gls{bc} \cite{Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hu2021b,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} to combine passive beamforming and backscatter modulation in the overall reflection pattern.
	In particular, \emph{symbol level precoding} maps the information symbols to the optimized \gls{ris} coefficient sets \cite{Liu2019d,Bereyhi2020}, \emph{overlay modulation} superposes the information symbols over a common auxiliary matrix \cite{Xu2020b,Zhang2021d,Hu2021b,Hua2022}, \emph{spatial modulation} switches between the reflection coefficient sets that maximize \gls{snr} at different receive antennas \cite{Basar2020,Ma2020a,Yuan2021}, and \emph{index modulation} employs dedicated reflection elements (resp. information elements) for passive beamforming (resp. backscatter modulation) \cite{Hu2021a}.
	However, those joint designs incur advanced hardware architecture and high optimization complexity.
	Relevant literatures also consider either \emph{Gaussian} codebook \cite{Guo2019b,Ding2020,Long2020a,Zhou2019a,Wu2021a,Xu2021a,Yang2021a,Hu2021b} that is impractical for low-power nodes, or finite \emph{equiprobable} inputs \cite{Yang2018,Liang2020,Han2021,Zhang2022,Liu2019d,Bereyhi2020,Xu2020b,Zhang2021d,Hua2022,Basar2020,Ma2020a,Yuan2021,Hu2021a} that does not fully exploit the \gls{csi} and signal properties of active-passive coexisting networks.
	% \end{subsection}

	% \begin{subsection}{Contributions}
	\emph{First,} we propose RIScatter as a novel scatter protocol that unifies \gls{bc} and \gls{ris} by adaptive reflection state (i.e., backscatter input) distribution design.
	As shown in Fig.~\subref*{fg:riscatter}, one or more passive RIScatter nodes ride over an active transmission and smartly scatters the incident wave to simultaneously engineer the primary channel and modulate its own information, while a co-located receiver cooperatively decodes both coexisting links.
	The reflection pattern over time is neither fully random like \gls{bc} nor fully deterministic like \gls{ris}, but semi-randomly chosen with the guidance of reflection state distribution, whose design accommodates the impact of information source, \gls{csi}, and \gls{qos} of both links.
	Such an adaptive backscatter channel coding boils down to the degenerate distribution of \gls{ris} when the primary link is prioritized, and outperforms the uniform distribution of line-coded \gls{bc} (by accounting the \gls{csi}) when the backscatter link is prioritized.
	Besides, when multiple RIScatter nodes are co-located, joint encoding can further increase the total backscatter rate.
	RIScatter exploits the transmitter-scatter-receiver path for simultaneous backscatter modulation and dynamic passive beamforming, and can smoothly transition in between under different \gls{qos}.
	However, such a novel scatter protocol encounters two major challenges, namely the cooperative receiver design and the adaptive backscatter input distribution design.
	This is the first paper to unify \gls{bc} and \gls{ris} from the perspective of reflection state distribution.

@@

	\emph{Fourth,} we provide numerical results to demonstrate the benefits of RIScatter and proposed algorithms.
	The observations include:
	1) adaptive reflection state distribution design can flexibly transition between backscatter modulation and passive beamforming;
	2) when the primary link is prioritized, input distribution becomes degenerate and RIScatter nodes coincide with discrete \gls{ris};
	3) when the backscatter link is prioritized, adaptive RIScatter encoding achieves higher backscatter rate than conventional line-coded \gls{bc} with equiprobable inputs;
	4) co-located RIScatter nodes can further leverage total backscatter rate by joint encoding;
	5) the proposed receiver maintains the passive beamforming benefit and provides comparable backscatter rate to \gls{sic}-based \gls{sr}, with re-encoding costs reduced to $1/N$ and no re-precoding/cancellation;
	6) it also supports scatter nodes with faster load-switching speed for potentially higher throughput;
	7) \gls{pgd} active beamformer effectively increases the primary (resp. backscatter) rate by boosting the receive \gls{snr} (resp. widening the energy gap under different reflection states), and can smoothly transition in between under different \gls{qos};
	8) distribution-aware backscatter detectors provide higher backscatter rate than the conventional \gls{ml} detector.
	% \end{subsection}

@@

		\def\svgwidth{0.4\columnwidth}

@@

	Due to double modulation, the primary equivalent channel is a function of backscatter symbol tuple
		\boldsymbol{h}_{\text{E}}^\mathsf{H}(x_{\mathcal{K}}) \triangleq \boldsymbol{h}_{\text{D}}^\mathsf{H} + \sum_{k} \alpha_k \boldsymbol{h}_{\text{C},k}^\mathsf{H} x_k = \boldsymbol{h}_{\text{D}}^\mathsf{H} + \boldsymbol{x}^\mathsf{H} \mathrm{diag}(\boldsymbol{\alpha}) \boldsymbol{H}_{\text{C}},

@@
